Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-28 paper/casestudies.tex
No directory for section, it breaks all nice features of emacs and bibligraphy
Figures

\section{Random Generators Synthesis} \label{sec:synthesis}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

>>>>

\section{Case Studies}

Intuitively, adding structure information to a random generator should provide
better results when verifying a system using random testing.
%
% For instance, by generating values that are no only syntactically but also
% semantically meaningful we would expect to reach deeper stages of our program
% pipeline other than merely input parsing/validation.
%
This section describes two case studies showing that considering structure
information when deriving generators consistenly produces better testing
results.

Instead of restricting our scope to Haskell, in this work we followed a broader
evaluation approach taken previously to compare the performance of other
state-of-the-art derivation techniques
\cite{grieco2017,DBLP:conf/haskell/MistaRH18}.
%
Concretely, we evaluate how including additional structure information when
generating a set of random test cases (often referred as a \emph{corpus})
affects the code coverage obtained when using them to test a given target
program.
%
For this purpose, we targeted two external programs which expect highly
structured inputs, namely \emph{GNU CLISP}---the GNU Common Lisp compiler, and
\emph{HTML Tidy}---a well known HTML refactoring and correction utility.
%
We remark that these applications are not written in Haskell.
%
However, there exist Haskell libraries defining ADTs encoding their input
structure, i.e., Lisp and HTML values respectively. These libraries are:
\emph{hs-zuramaru}, implementing an embedded Lisp intepreter for a small subset
of this programming language, and \emph{html}, defining a combinator library for
constructing HTML values.
%
These libraries also come with serialization functions to map Haskell values
into corresponding test case files.

We started by compiling instrumented versions of the target programs in a way
that they also return the execution path followed in the source code every time
we run them with a given input test case.
%
This let us distinguish the amount of different execution paths that a randomly
generated corpus can trigger.
%
On the other side, we used the ADTs defined on the chosen libraries to derive
random generators using \dragen and \dragenp, including structure information
extracted from the library's codebase in the case of the latter.
%
Then, we evaluated the code coverage triggered by independent, randomly
generated corpora of different sizes varying from 100 to 1000 test cases each.
%
In order to remove any external bias, we derived random generators optimized to
follow a uniform distribution of constructors (and pattern matchings or function
calls in the case \dragenp), and carefully adjusted their generation sizes to
match the average test case size in bytes.
%
This way, any noticeable difference in the code coverage can be attributed to
the present (or lack thereof) structure information when generating the test
cases.
%
Additionally, to achieve statistical significance we repeated each experiment 30
times with independently generated sets of random test cases.

\begin{figure*}[t]
  \centering
  \input{tikz/lisp.tex}
  \hspace{5pt}%
  \input{tikz/html.tex}
  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
    \dragenp (\ref{exp:dragenp}). }
  \label{fig:coverage}
\end{figure*}

Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
triggered for each permutation of corpus size and derivation tool, including
error bars indicating the standard error of the mean on each case.
%
We proceed to describe each case study in detail.

\paragraph{Generating Lisp code using pattern matching information}

In this first case study we wanted to evaluate the observed code coverage
differences when considering structure information present on functions pattern
matchings.

Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
represented as plain strings; and literal values like booleans or integers.
%
In order to interpret Lisp programs, this unified representation of data and
code requires this library to pattern match against common patterns like
let-bindings, if-then-else expressions and arithmetic operators among others.
%
In particular, each one of these patterns match a against special symbol of the
Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
sub-expressions.

In this case study we extracted the structure information from the pattern
matchings present on this library and included it into the generation
specification of our random Lisp values, which were generated by randomly
picking from a total of 6 data constructors and 8 different pattern matchings.
%
By doing this we obtained a consistent code coverage improvement of
approximately $4\%$ using \dragenp with respect to the one obtained with
\dragen (see Figure \ref{fig:coverage} (a)).
%
\todo[inline, author=AM]{Shall we say anything justifing this little
  improvement?}

\paragraph{Generating HTML pages using abstract interface information}

For our second case study, we wanted to evaluate how including structural
information coming from abstract interfaces when generating random HTML values
might improve the testing performance.

The library we used for this purpose represents HTML values very much in the
same way as we exemplify in Section \ref{sec:randomtesting}, i.e., defining a
small set of general constructions representing plain text and tags---altough
this library also supports HTML tag attributes as well.
%
Then, this representation is extended with a large abstract interface consisting
of combinators representing common HTML tags and tag attributes---equivalent to
the combinators |div|, |bold| and |hr| illustrated before.

In this case study we included the structure information present on the abstract
interface of this library into the generation specification of random HTML
values, resulting in a generation process that randomly picked among 4 data
constructors and 163 abstract combinators.
%
With this large amount of additional structure information, we observed an
increase of up to $83\%$ in the code coverage obtained with \dragenp with
respect to the one observed with \dragen (see Figure \ref{fig:coverage} (b)).

A manual inspection of the corpora generated with each tool revealed us that, in
general terms, the test cases generated with \dragen rarely represent
syntactically correct HTML values, consisting to a large extent of random
strings within and between HTML tag delimeters (|"<"|, |">"| and |"/>"|).
%
On the other hand, test cases generated with \dragenp encode much more
interesting structural information, being mostly syntactically correct.
%
We found that, in many cases, the test cases generated with \dragenp were
parsed, analysed and reported as valid HTML values by the target application.

With these results we are confident that including the structural information
present on the user codebase improves the testing performance.
%
We consider that our approach is particularly useful when the data types
encoding the shape of our data are vague or not sufficiently structured to be
used to derive powerful random generators with the common type-driven
derivation techniques.
